### 2015-12-16 paper/paper.tex
added start of final paper

\documentclass{article} % For LaTeX2e
\usepackage{nips15submit_e,times}
\usepackage{hyperref}
\usepackage{url}
%\documentstyle[nips14submit_09,times,art10]{article} % For LaTeX 2.09
\newcommand\BibTeX{B{\sc ib}\TeX}

\title{LSTM for Sentiment Analysis on Twitter}

\author{
Trapit Bansal
\And
Kate Silverstein
\And
Jun Wang
}

%% \author{
%% David S.~Hippocampus\thanks{ Use footnote for providing further information
%% about author (webpage, alternative address)---\emph{not} for acknowledging
%% funding agencies.} \\
%% Department of Computer Science\\
%% Cranberry-Lemon University\\
%% Pittsburgh, PA 15213 \\
%% \texttt{hippo@cs.cranberry-lemon.edu} \\
%% \And
%% Coauthor \\
%% Affiliation \\
%% Address \\
%% \texttt{email} \\
%% \AND
%% Coauthor \\
%% Affiliation \\
%% Address \\
%% \texttt{email} \\
%% \And
%% Coauthor \\
%% Affiliation \\
%% Address \\
%% \texttt{email} \\
%% \And
%% Coauthor \\
%% Affiliation \\
%% Address \\
%% \texttt{email} \\
%% (if needed)\\
%% }

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\nipsfinalcopy % Uncomment for camera-ready version

\begin{document}

\maketitle

\begin{abstract}
abstract goes here
\end{abstract}

\section{Introduction}

We use LSTM \cite{hochreiter1997long} to do stuff

\section{Related Work}

\section{Model}

\section{Experiments}
See Table ~\ref{sample-table} for awesome results

\begin{table}[h]
\caption{Sample table title}
\label{sample-table}
\begin{center}
\begin{tabular}{ll}
\multicolumn{1}{c}{\bf PART}  &\multicolumn{1}{c}{\bf DESCRIPTION}
\\ \hline \\
Dendrite         &Input terminal \\
Axon             &Output terminal \\
Soma             &Cell body (contains cell nucleus) \\
\end{tabular}
\end{center}
\end{table}

This is a figure:

\begin{figure}[h]
\begin{center}
%\framebox[4.0in]{$\;$}
\fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
\end{center}
\caption{Sample figure caption.}
\end{figure}

\section{Conclusion}

\subsubsection*{Acknowledgments}

\subsubsection*{References}

\bibliography{paper}
\bibliographystyle{plain}
%% References follow the acknowledgments. Use unnumbered third level heading for
%% the references. Any choice of citation style is acceptable as long as you are
%% consistent. It is permissible to reduce the font size to `small' (9-point) 
%% when listing the references. {\bf Remember that this year you can use
%% a ninth page as long as it contains \emph{only} cited references.}

%% \small{
%% [1] Alexander, J.A. \& Mozer, M.C. (1995) Template-based algorithms
%% for connectionist rule extraction. In G. Tesauro, D. S. Touretzky
%% and T.K. Leen (eds.), {\it Advances in Neural Information Processing
%% Systems 7}, pp. 609-616. Cambridge, MA: MIT Press.

%% [2] Bower, J.M. \& Beeman, D. (1995) {\it The Book of GENESIS: Exploring
%% Realistic Neural Models with the GEneral NEural SImulation System.}
%% New York: TELOS/Springer-Verlag.

%% [3] Hasselmo, M.E., Schnell, E. \& Barkai, E. (1995) Dynamics of learning
%% and recall at excitatory recurrent synapses and cholinergic modulation
%% in rat hippocampal region CA3. {\it Journal of Neuroscience}
%% {\bf 15}(7):5249-5262.
%% }

\end{document}

### 2015-12-18 semeval/lstm_chars.py
write learning curve data to separate file

@@ -3,7 +3,7 @@ import cPickle
 import os
 import time
 import argparse
-import codecs
+from collections import OrderedDict
 
 import numpy as np
 import theano
@@ -17,6 +17,7 @@ MAXLEN = 140
 
 
 def build_model(vmap,
+                log,
                 nclasses=2,
                 embedding_dim=50,
                 nhidden=256,
@@ -24,22 +25,14 @@ def build_model(vmap,
                 invar=None,
                 maskvar=None,
                 bidirectional=True,
-                pool=True,
+                pool='mean',
                 grad_clip=100,
                 maxlen=MAXLEN):
 
+    net = OrderedDict()
+
     V = len(vmap)
     W = lasagne.init.Normal()
-
-    # Input Layer
-    l_in = layer.InputLayer((batchsize, maxlen), input_var=invar)
-    l_mask = layer.InputLayer((batchsize, maxlen), input_var=maskvar)
-    ASSUME = {l_in: (200, 140), l_mask: (200, 140)}
-
-    # Embedding Layer
-    l_emb = layer.EmbeddingLayer(l_in, input_size=V, output_size=embedding_dim, W=W)
-    print 'Embedding Layer'
-    print 'output:', get_output_shape(l_emb, ASSUME)
 
     gate_params = layer.recurrent.Gate(
         W_in=lasagne.init.Orthogonal(),
@@ -55,30 +48,29 @@ def build_model(vmap,
         nonlinearity=lasagne.nonlinearities.tanh
     )
 
-    l_fwd = layer.LSTMLayer(
-        l_emb,
+    net['input'] = layer.InputLayer((batchsize, maxlen), input_var=invar)
+    net['mask'] = layer.InputLayer((batchsize, maxlen), input_var=maskvar)
+    ASSUME = {net['input']: (200, 140), net['mask']: (200, 140)}
+    net['emb'] = layer.EmbeddingLayer(net['input'], input_size=V, output_size=embedding_dim, W=W)
+    net['fwd1'] = layer.LSTMLayer(
+        net['emb'],
         num_units=nhidden,
         grad_clipping=grad_clip,
         nonlinearity=lasagne.nonlinearities.tanh,
-        mask_input=l_mask,
+        mask_input=net['mask'],
         ingate=gate_params,
         forgetgate=gate_params,
         cell=cell_params,
         outgate=gate_params,
         learn_init=True
     )
-
-    print 'Forward LSTM'
[…]

### 2015-12-18 semeval/lstm_chars.py
add fxn to test models

@@ -5,6 +5,8 @@ import time
 import argparse
 from collections import OrderedDict
 
+from evaluate import ConfusionMatrix
+
 import numpy as np
 import theano
 import theano.tensor as T
@@ -172,7 +174,7 @@ def learn_model(train_path,
     print "Validation size", X_val.shape[0]
     print "Test size", X_test.shape[0]
 
-    V = len(vmap) 
+    V = len(vmap)
     n_classes = len(set(y_train))
     print "Vocab size:", V
     print "Number of classes", n_classes
@@ -412,10 +414,51 @@ def load_dataset(trainfile, testfile, vocabfile, devfile=None, pad_with=0):
     return train, dev, test, vocab_shift
 
 
+def test_model(model_path,
+               train_path,
+               val_path=None,
+               test_path=None,
+               vocab_file=None,
+               label_file=None,
+               log_path="",
+               embeddings_file=None):
 
+    timestamp = time.strftime('%m%d%Y_%H%M%S')
+    log_file = open(log_path + "/test_results_" + timestamp, "w+")
+    print "Loading Dataset"
+    _, dev, test, vmap = load_dataset(train_path, test_path, vocab_file, devfile=val_path)
+    y_val, X_val = dev
+    y_test, X_test = test
+    n_classes = len(set(y_val))
+    pad_char = u'♥'
+    vmap[pad_char] = 0
 
+    # Initialize theano variables for input and output
+    X = T.imatrix('X')
+    M = T.matrix('M')
+    y = T.ivector('y')
 
+    # Construct network
+    print "Building Model"
+    log_file.write('model file: %s' % model_path)
+    network = build_model(vmap, log_file, n_classes, invar=X, maskvar=M)
+    read_model_data(network, model_path)
 
+    # need to switch off droput while testing
+    test_output = lasagne.layers.get_output(network['softmax'], deterministic=True)
+    val_cost_fn = lasagne.objectives.categorical_crossentropy(test_output, y).mean()
+    preds = T.argmax(test_output, axis=1)
+    val_acc_fn = T.mean(T.eq(preds, y), dtype=theano.config.floatX)
+    val_fn = theano.function([X, M, y], [val_cost_fn, val_acc_fn, preds], allow_input_downcast=True)
+
+    val_loss, val_acc, val_pred = val_fn(X_val[:, :, 0], X_val[:, :, 1], y_val)
+
[…]

### 2015-12-18 semeval/hparams.py
some code cleanup

@@ -1,0 +1,52 @@
+class HParams(object):
+    def __init__(self,
+                 nepochs=30,
+                 batchsize=512,
+                 learning_rate=0.1,
+                 bidirectional=True,
+                 nhidden=256,
+                 embedding_dim=50,
+                 pool='mean',
+                 grad_clip=100,
+                 optimizer='adam'):
+        
+        self.nepochs = nepochs
+        self.batchsize = batchsize
+        self.learning_rate = learning_rate
+        self.bidirectional = bool(bidirectional)
+        self.nhidden = nhidden
+        self.embedding_dim = embedding_dim
+        self.pool = pool
+        self.grad_clip = grad_clip
+        self.optimizer = optimizer
+    
+    def parse_args(self, args):        
+        self.nepochs = args.nepochs
+        self.batchsize = args.batchsize
+        self.learning_rate = args.learning_rate
+        self.bidirectional = bool(args.bidirectional)
+        self.nhidden = args.nhidden
+        self.embedding_dim = args.embedding_dim
+        self.pool = args.pool
+        self.grad_clip = args.grad_clip
+        self.optimizer = args.optimizer
+        
+    def __repr__(self):
+        hps = {}
+        hps['nepochs'] = self.nepochs
+        hps['batchsize'] = self.batchsize
+        hps['learning_rate'] = self.learning_rate
+        hps['bidirectional'] = self.bidirectional
+        hps['nhidden'] = self.nhidden
+        hps['embedding_dim'] = self.embedding_dim
+        hps['pool'] = self.pool
+        hps['grad_clip'] = self.grad_clip
+        hps['optimizer'] = self.optimizer
+        srt = sorted(hps.items(), key=lambda x: x[0])
+        head = '========== HPARAMS ========='
+        string = head + '\n'
+        for k, v in srt:
+            string += '%s: %s\n' % (k, v)
+        end = ''.join(['=']*len(head))
+        string += end + '\n'
+        return string

### 2015-12-18 semeval/lstm_chars.py

@@ -4,32 +4,35 @@ import os
 import time
 import argparse
 from collections import OrderedDict
-
-from evaluate import ConfusionMatrix
-
 import numpy as np
+
 import theano
 import theano.tensor as T
 import lasagne
 import lasagne.layers as layer
 from lasagne.layers import get_output_shape
 
+from evaluate import ConfusionMatrix
+from hparams import HParams
 
 MAXLEN = 140
+SEED = 1234
 
 
-def build_model(vmap,
+def build_model(hyparams,
+                vmap,
                 log,
                 nclasses=2,
-                embedding_dim=50,
-                nhidden=256,
                 batchsize=None,
                 invar=None,
                 maskvar=None,
-                bidirectional=True,
-                pool='mean',
-                grad_clip=100,
                 maxlen=MAXLEN):
+
+    embedding_dim = hyparams.embedding_dim
+    nhidden = hyparams.nhidden
+    bidirectional = hyparams.bidirectional
+    pool = hyparams.pool
+    grad_clip = hyparams.grad_clip
 
     net = OrderedDict()
 
@@ -126,13 +129,14 @@ def build_model(vmap,
     return net
 
 
-def iterate_minibatches(inputs, targets, batchsize, shuffle=False):
+def iterate_minibatches(inputs, targets, batchsize, rng=None, shuffle=False):
     ''' Taken from the mnist.py example of Lasagne'''
     # print inputs.shape, targets.size
     assert inputs.shape[0] == targets.size
     if shuffle:
+        assert rng is not None
         indices = np.arange(inputs.shape[0])
-        np.random.shuffle(indices)
+        rng.shuffle(indices)
     for start_idx in range(0, inputs.shape[0] - batchsize + 1, batchsize):
         if shuffle:
             excerpt = indices[start_idx:start_idx + batchsize]
@@ -141,27 +145,25 @@ def iterate_minibatches(inputs, targets, batchsize, shuffle=False):
         yield inputs[excerpt], targets[excerpt]
 
 
-def learn_model(train_path,
+def learn_model(hyparams,
+                train_path,
                 val_path=None,
                 test_path=None,
-                max_norm=5,
-                num_epochs=5,
-                batchsize=10,  # 64,
-                learn_rate=0.1,
                 vocab_file=None,
-                val_ratio=0.1,
                 log_path="",
+                model_file=None,
[…]

### 2015-12-18 train-lstm-chars.sh

@@ -1,17 +1,24 @@
 #!/usr/bin/env bash
 
 root="$PWD"
-dataRoot="$root/CHAR_DATA/big_ascii"
+dataRoot="/home/kate/F15/CHAR_DATA"
 
-infile="$dataRoot/tweets.chars.tsv.small"
-vocab="$root/CHAR_DATA/ascii_vocab.pkl"
-testfile="$dataRoot/tweets.chars.tsv.test"
+infile="$dataRoot/big_ascii/tweets.chars.tsv.small"
+testfile="$dataRoot/big_ascii/tweets.chars.tsv.test"
+vocab="$dataRoot/ascii_vocab.pkl"
+labels="$dataRoot/labels.pkl"
 
-logdir="$root/char_lstm_result_ascii"
+logdir="$root/debug_ascii"
 mkdir -pv $logdir
 
 THEANO_FLAGS=device=cpu,floatX=float32 python $root/semeval/lstm_chars.py \
 --tweet-file $infile \
 --vocab $vocab \
 --log-path $logdir \
---test-file $testfile
+--test-file $testfile \
+--batchsize "100" \
+--nepochs "1" \
+--model-file "$root/bidir_lstm_ascii" \
+--results-file "$logdir/bidir_ascii_results.txt" \
+--label-file $labels
+

### 2015-12-19 paper-Jun/paper.tex
Update paper (intro & background)

--- 2015-12-16 paper/paper.tex
@@
 \section{Introduction}
+% why analysis sentiment expecially twitter
+In the last fewer year, microblogging has become a very popular communication tool in people's social life.
+On microblogging platforms, such as Twitter\footnote{\url{https://twitter.com/}} and Facebook\footnote{\url{https://www.facebook.com/}}, a diverge range of people are attracted to post short sentences, images, and video links to share life issues and opinions.
+This popularity results in enormous amount of information coving a wide range of topics from celebrities to opinions about the brands, products, politicians and social events. 
+Such data is a valuable and efficient source for marketing and social studies, and therefore sentiment analysis on these data have obtained special interest. 
+For example, manufacturing companies may want to know how people like their product (or service), what would people prefer.
 
-We use LSTM \cite{hochreiter1997long} to do stuff
+We restrict our interest on Twitter sentiment analysis for following reasons: 
+People are posting on various topics on Twitter, thus it is a valuable source of data; Twitter is one of most popular microblogging platform and the number of messages are growing everyday.
+The collected corp can be arbitrary large; Users of Twitter have various background and have different language conventions It is possible to collect tweets of large diversity in both topics and language model.
+%Some typical examples of tweets are shown in Table \ref{tweet-example}. 
+%
+%\begin{table}[h]
+%	\caption{Examples of typical tweets}
+%	\label{tweet-example}
+%	\begin{center}
+%		\begin{tabular}{l}
+%			\hline
+%			Edelman11 Cant. Wait. \#TheForceAwakens \#starpats \\\hline
+%			JeriLRyan SQUEEEEEEEEEE!!!!! Thank you, @Sphero!!! (And @bonniegrrl, you rock!!)   \\\hline
+%			@appcompany Why haven’t you updated your app? Apple released a new iOS and your app now doesn’t work.\\ \hline
+%		\end{tabular}
+%	\end{center}
+%\end{table}
+
+Compared with traditional blogs, tweets have some unique characteristics: 1.
+Tweets are short in length.
+There is a limitation of $140$ words for each tweet; 2.
+The language used in tweets is very information with misspelling, creative spelling,  new words, slang, and URLs; 3.
+Emoticons and hashtags are frequently used.
+
+% what we did
+In this paper, we propose a bi-direction LSTM (Long Short -Term Memory)\cite{hochreiter1997long} method for sentiment analysis.
+We tried both word-level and character-level features on the bi-direction LSTM model and compare the results with DCNN (Dynamic Convolutional Neural Network) \cite{kalchbrenner2014convolutional}.
+We show that the accuracy of sentiment analysis ... (Please revise this part.)
+
+We train our model on $1.6$M distantly-supervised tweets collected by Go et.
+al.
+\cite{go2009twitter}, and evaluate the result on SemEval-2016 Task 4.
+Currently we focus on polarity classification, and plan to apply our model on 5-point scale classification in the future.
+
+% the structure of the paper
+In the remaining of this paper, we first introduce the related work, the dataset, and the evaluation methodology.
+We then describe the model we proposed.
+Finally, we discuss the experiment results and point to possible directions for future research.
+(Please revise this part.)
 
 \section{Related Work}
+Twitter sentiment analysis is increasingly drawing attention of researchers in recent years. 
+Given the length limitations on tweets, sentiment analysis of tweets is often considered similar to sentence-level sentiment analysis \cite{kouloumpis2011twitter}.
+However, phrase and sentence level approaches can hardly define the sentiment of some specific topic.
+Considering opinions adhering on different topics, Wang et.
+al.\cite{wang2011topic} proposed a hashtag-level sentiment classification method  to generate the overall sentiment polarity for a given hashtag.
+Recently, following the work of \cite{mikolov2013efficient} some researchers employed neural network to implement sentiment classification. 
+For example, Kim \cite{kim2014convolutional} adopted convolutional neural networks to learn sentiment-bearing sentence vectors, Mikolov et al.\cite{mikolov2013distributed} proposed Paragraph vector which outperformed bag-of-words model for sentiment analysis, and Tang et.
+al.
+\cite{tang2014learning} used ConvNets to learn sentiment specific word embedding (SSWE), which encodes sentiment information in the continuous
+representation of words.
+Furthermore, Kalchbrenner \cite{kalchbrenner2014convolutional} proposed a Dynamic Convolutional Neural Network (DCNN) which uses dynamic k-max pooling, a global pooling operation over linear sequences.
+Instead of directly applying ConvNets to embeddings of words, \cite{zhang2015character} apply the network only on characters.
+They showed that the deep ConvNets do not require knowledge of words and thus can work for different languages.
 

### 2015-12-19 semeval/hparams.py
add option to init second layer with identity matrix

@@ -8,7 +8,8 @@ class HParams(object):
                  embedding_dim=50,
                  pool='mean',
                  grad_clip=100,
-                 optimizer='adam'):
+                 optimizer='adam',
+                 init='random'):
         
         self.nepochs = nepochs
         self.batchsize = batchsize
@@ -19,6 +20,7 @@ class HParams(object):
         self.pool = pool
         self.grad_clip = grad_clip
         self.optimizer = optimizer
+        self.init = init
     
     def parse_args(self, args):        
         self.nepochs = args.nepochs
@@ -30,6 +32,7 @@ class HParams(object):
         self.pool = args.pool
         self.grad_clip = args.grad_clip
         self.optimizer = args.optimizer
+        self.init = args.init
         
     def __repr__(self):
         hps = {}

### 2015-12-19 semeval/lstm_chars.py

@@ -33,6 +33,7 @@ def build_model(hyparams,
     bidirectional = hyparams.bidirectional
     pool = hyparams.pool
     grad_clip = hyparams.grad_clip
+    init = hyparams.init
 
     net = OrderedDict()
 
@@ -44,7 +45,6 @@ def build_model(hyparams,
         W_hid=lasagne.init.Orthogonal(),
         b=lasagne.init.Constant(0.)
     )
-
     cell_params = layer.recurrent.Gate(
         W_in=lasagne.init.Orthogonal(),
         W_hid=lasagne.init.Orthogonal(),
@@ -96,19 +96,46 @@ def build_model(hyparams,
     else:
         net['pool'] = layer.ConcatLayer([net['fwd1']])
     net['dropout1'] = layer.DropoutLayer(net['pool'], p=0.5)
-    net['fwd2'] = layer.LSTMLayer(
-        net['dropout1'],
-        num_units=nhidden,
-        grad_clipping=grad_clip,
-        nonlinearity=lasagne.nonlinearities.tanh,
-        mask_input=net['mask'],
-        ingate=gate_params,
-        forgetgate=gate_params,
-        cell=cell_params,
-        outgate=gate_params,
-        learn_init=True,
-        only_return_final=True
-    )
+    if init == 'identity':
+        gate_params2 = layer.recurrent.Gate(
+            W_in=np.eye(nhidden, dtype=np.float32),
+            W_hid=np.eye(nhidden, dtype=np.float32),
+            b=lasagne.init.Constant(0.)
+        )
+        cell_params2 = layer.recurrent.Gate(
+            W_in=np.eye(nhidden, dtype=np.float32),
+            W_hid=np.eye(nhidden, dtype=np.float32),
+            W_cell=None,
+            b=lasagne.init.Constant(0.),
+            nonlinearity=lasagne.nonlinearities.rectify
+        )
+        net['fwd2'] = layer.LSTMLayer(
+            net['dropout1'],
+            num_units=nhidden,
+            grad_clipping=grad_clip,
+            nonlinearity=lasagne.nonlinearities.tanh,
+            mask_input=net['mask'],
+            ingate=gate_params2,
+            forgetgate=gate_params2,
+            cell=cell_params2,
+            outgate=gate_params2,
+            learn_init=True,
+            only_return_final=True
+        )
+    else:
+        net['fwd2'] = layer.LSTMLayer(
+            net['dropout1'],
+            num_units=nhidden,
+            grad_clipping=grad_clip,
[…]

### 2015-12-19 semeval/hparams.py
add option to init second layer with identity matrix

@@ -45,6 +45,7 @@ class HParams(object):
         hps['pool'] = self.pool
         hps['grad_clip'] = self.grad_clip
         hps['optimizer'] = self.optimizer
+        hps['init'] = self.init
         srt = sorted(hps.items(), key=lambda x: x[0])
         head = '========== HPARAMS ========='
         string = head + '\n'

### 2015-12-19 semeval/lstm_chars.py

@@ -295,7 +295,8 @@ def learn_model(hyparams,
 
     print "Starting Training"
     nbatches = X_train.shape[0] / batchsize
-    valfreq = max(int(nbatches / 32.), 2)  # evaluate every [valfreq] minibatches
+    valfreq = max(int(nbatches / 8.), 2)  # evaluate every [valfreq] minibatches
+    print 'validating every', valfreq, 'batches'
 
     begin_time = time.time()
     best_val_acc = -np.inf

### 2015-12-19 paper-Jun/paper.tex
Revised intro and related work

--- 2015-12-16 paper/paper.tex
@@
 \section{Introduction}
+% why analysis sentiment expecially twitter
+In the last few years, microblogging has become a very popular communication tool in people's social life.
+On microblogging platforms, such as Twitter\footnote{\url{https://twitter.com/}} and Facebook\footnote{\url{https://www.facebook.com/}}, a diverse range of people are attracted to post short sentences, images, and video links to share life issues and opinions.
+This popularity results in enormous amount of information covering a wide range of topics on the brands, products, politics and social events. 
+Such data is a valuable and efficient source for marketing and social studies. 
+Sentiment analysis on these data, which determines the attitude of a user with respect to some topic or product, will provide convincing information.
+For example, it may help manufacturing companies to know how people like their product (or service), what would people prefer.
+Therefore sentiment analysis on microblogging has obtained special interest.
 
-We use LSTM \cite{hochreiter1997long} to do stuff
+In this paper, we focus on using sentiment analysis on Twitter. 
+We restrict our interest on Twitter for the following reasons: 
+People are posting on various topics on Twitter, thus it is a valuable source of data; Twitter is one of the most popular microblogging platform and the number of messages are growing everyday.
+The collected corp can be arbitrary large; Users of Twitter have various background and have different language conventions.
+It is possible to collect tweets of wide diversity in both topics and language models.
+
+Although there exist plenty of work on text classification, some unique characteristics of tweets present special challenges for sentiment analysis:
+1.
+Tweets are short in length.
+There is a limitation of $140$ words for each tweet; 2.
+The language used in tweets is very informal with misspelling, creative spelling,  new words, slangs, and URLs; 3.
+Emotions and hashtags are frequently used.
+
+% what we did
+In this paper, we propose a bi-directional Long Short-Term Memory (LSTM) method for sentiment analysis.
+We tried both word-level and character-level features on the bi-directional LSTM model and compare the results with Dynamic Convolutional Neural Network (DCNN) \cite{kalchbrenner2014convolutional}.
+We show that the accuracy of sentiment analysis ... (Please revise this part.)
+
+We train our model on $1.6$ M distantly-supervised tweets collected by Go et.
+al.
+\cite{go2009twitter}, and evaluate the results on SemEval-2016 Task 4\footnote{\url{http://alt.qcri.org/semeval2016/task4/}}.
+Currently we focus on polarity classification, and plan to apply our model on 5-point scale classification in the future.
+
+% the structure of the paper
+In the remaining of this paper, we first introduce the related work, the dataset, and the evaluation methodology.
+We then describe the model we proposed.
+Finally, we discuss the experiment results and point out possible directions for future research.
+(Please revise this part.)
 
 \section{Related Work}
+Twitter sentiment analysis is increasingly drawing attention of researchers in recent years. 
+Given the length limitations on tweets, sentiment analysis of tweets is often considered similar to sentence-level sentiment analysis \cite{kouloumpis2011twitter}.
+However, phrase and sentence level approaches can hardly define the sentiment of some specific topics.
+Considering opinions adhering on different topics, Wang et.
+al.
+\cite{wang2011topic} proposed a hashtag-level sentiment classification method  to generate the overall sentiment polarity for a given hashtag.
+Recently, following the work of \cite{mikolov2013efficient} some researchers used neural network to implement sentiment classification. 
+For example, Kim \cite{kim2014convolutional} adopted convolutional neural networks to learn sentiment-bearing sentence vectors, Mikolov et al.
+\cite{mikolov2013distributed} proposed Paragraph vector which outperformed bag-of-words model for sentiment analysis, and Tang et.
+al.
+\cite{tang2014learning} used ConvNets to learn sentiment specific word embedding (SSWE), which encodes sentiment information in the continuous
+representation of words.
+Furthermore, Kalchbrenner \cite{kalchbrenner2014convolutional} proposed a Dynamic Convolutional Neural Network (DCNN) which uses dynamic k-max pooling, a global pooling operation over linear sequences.
+Instead of directly applying ConvNets to embeddings of words, \cite{zhang2015character} applies the network only on characters.
+They showed that the deep ConvNets does not require knowledge of words and thus can work for different languages.
+LSTM \cite{hochreiter1997long} is another state-of-the-art semantic composition models for sentiment classification \cite{li2015tree}.
+Similar to DCNN, it also learns fixed-length vectors for sentences of varying length, captures words order in a sentence and does not depend on external dependency or constituency parse results.
 

### 2015-12-19 paper-Jun/paper.tex
Revised intro and related work

@@
 Such data is a valuable and efficient source for marketing and social studies. 
-Sentiment analysis on these data, which determines the attitude of a user with respect to some topic or product, will provide convincing information.
+Sentiment analysis on microblogging has obtained special interest, because it determines the attitude of a user with respect to some topic or product and thus provides provide convincing information
 For example, it may help manufacturing companies to know how people like their product (or service), what would people prefer.
-Therefore sentiment analysis on microblogging has obtained special interest.
 

### 2015-12-20 paper/paper.tex
add some refs
add some figs
fill in some results, add a fig

@@
 \usepackage{hyperref}
+\usepackage{graphicx}
 \usepackage{url}
@@
 
+Dynamic convolutional neural net for sentences \cite{kalchbrenner2014convolutional}
+
+sentiment specific word embeddings \cite{tang2014learning}
+
+character-lvl CNN for text classification \cite{zhang2015character}
+
 \section{Model}
@@
 \section{Experiments}
-See Table ~\ref{sample-table} for awesome results
 
-\begin{table}[h]
-\caption{Sample table title}
-\label{sample-table}
-\begin{center}
-\begin{tabular}{ll}
-\multicolumn{1}{c}{\bf PART}  &\multicolumn{1}{c}{\bf DESCRIPTION}
-\\ \hline \\
-Dendrite         &Input terminal \\
-Axon             &Output terminal \\
-Soma             &Cell body (contains cell nucleus) \\
+\subsection{Datasets}
+
+\subsection{Results}
+
+\begin{table}[h!]
+\centering
+\caption{Accuracy across LSTMs}
+\label{char-models}
+\begin{tabular}{|l|ll|}
+\hline
+              & 1.6M           & semeval \\
+\hline
+ascii/rnd/50  & \textbf{83.84} & 82.08   \\
+ascii/rnd/200 & 82.45          & 83.53   \\
+ascii/eye/50  & 77.44          &         \\
+utf8          & 81.34          & 82.34   \\
+word/rnd      & 81.85          & 78.07   \\
+word/sswe     & 83.24          & 79.27   \\
+\hline
 \end{tabular}
-\end{center}
 \end{table}
-
-This is a figure:
 
@@
 %\framebox[4.0in]{$\;$}
-\fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
+\includegraphics[width=0.95\textwidth]{figs/cool}
+%\fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
 \end{center}
 \caption{Sample figure caption.}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+%\framebox[4.0in]{$\;$}
+\includegraphics[width=0.95\textwidth]{figs/puppies}
+%\fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
+\end{center}
+\caption{Sample figure caption.}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+%\framebox[4.0in]{$\;$}
+\includegraphics[width=0.95\textwidth]{figs/cool}
+%\fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
+\end{center}
+\caption{Sample figure caption.}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+%\framebox[4.0in]{$\;$}
+\includegraphics[width=0.95\textwidth]{figs/dentist}
+%\fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
+\end{center}
+\caption{Confidence in ``negative'' prediction over time}
 \end{figure}

### 2015-12-20 paper/paper.tex
fill in some corpus stats

@@
 
-\section{Experiments}
+\section{Results}
 
 \subsection{Datasets}
+We conduct our experiments on two datasets: the latest benchmark dataset for SemEval 2016 and the dataset provided by \cite{go2009twitter}.
+In the latter dataset, the training set consists of $1.6$ million weakly-supervised tweets collected during $2009$, and the test set is hand-labeled.
+In the experiments presented below, we first train our models on the Go dataset, then re-train the model parameters on the smaller, fully-supervised SemEval dataset.
 
-\subsection{Results}
+\begin{table}[h!]
+\centering
+\caption{Data size and label distribution}
+\label{my-label}
+\begin{tabular}{|lllll|}
+\hline
+      & \multicolumn{2}{l}{Go et. al. (1.6M)} & \multicolumn{2}{l}{SemEval2016} \\
+      & neg               & pos               & neg            & pos            \\
+\hline
+train & 800000            & 800000            & 781            & 2805           \\
+dev   & -                 & -                 & 358            & 766            \\
+test  & 177               & 182               & 286            & 886           \\
+\hline
+\end{tabular}
+\end{table}
+
+\subsection{Experiments}
+We compare the performance of character-level models against word-level models.
+For the former, we compare across character sets (``utf8'' and ``ascii''), parameter initializations (``rnd'' and ``eye''), and embedding dimension sizes ($50$ and $200$).
+The vocabulary for the ``utf8'' setting consisted of $1949$ characters, and for ``ascii'' consisted of $93$ characters.
+We initialize model parameters randomly at all levels in the ``rnd'' setting whereas in the ``eye'' setting, we initialize the second-level LSTM cell and gate parameters using the identity matrix.
+
+For the latter, we compare initializing the word embeddings using random vectors (``word/rnd'') against initialization using the embeddings provided in \cite{tang2014learning} (``word/sswe'').
 
@@
 \hline
-              & 1.6M           & semeval \\
+              & 1.6M (acc)          & semeval (acc) \\
 \hline
 ascii/rnd/50  & \textbf{83.84} & 82.08   \\
-ascii/rnd/200 & 82.45          & 83.53   \\
-ascii/eye/50  & 77.44          &         \\
+ascii/rnd/200 & 82.45          & \textbf{84.13}   \\
+ascii/eye/50  & 77.44          & 79.18   \\
 utf8          & 81.34          & 82.34   \\
@@
 \end{table}
+
+\subsection{Analysis}
 

### 2015-12-20 paper-Jun/paper.tex
done with related work

@@
 \usepackage{url}
+\usepackage{graphicx}
+\usepackage{amsmath}
 %\documentstyle[nips14submit_09,times,art10]{article} % For LaTeX 2.09
@@
 In this paper, we focus on using sentiment analysis on Twitter. 
-We restrict our interest on Twitter for the following reasons: 
-People are posting on various topics on Twitter, thus it is a valuable source of data; Twitter is one of the most popular microblogging platform and the number of messages are growing everyday.
-The collected corp can be arbitrary large; Users of Twitter have various background and have different language conventions.
+Twitter is an extremely popular microblogging platform which allows people to post messages of up to 140 characters.
+Because of the short nature of tweets, people often post twitter messages (called Tweets) frequently while attending events like product launches, movie premiers, music concerts or just to express their opinion on a trending or current topic.
+As such, they can be a valuable source of public opinion or feedback \cite{o2010tweets, bollen2011twitter, bollen2009modeling}.
+Realizing this importance, analyzing sentiment of Tweets has been a recurring task in the SemEval \footnote{\url{http://alt.qcri.org/semeval2016/task4/}} competitions.
+
+%We restrict our interest on Twitter for the following reasons: 
+%People are posting on various topics on Twitter, thus it is a valuable source of data; Twitter is one of the most popular microblogging platform and the number of messages are growing everyday.
+The collected corpus can be arbitrary large.
+Users of Twitter have various background and have different language conventions.
 It is possible to collect tweets of wide diversity in both topics and language models.
@@
 
-\section{Model}
+\subsection{Dynamic Convolutional Neural Networks (DCNN)}
+We briefly review the architecture of DCNN \cite{kalchbrenner2014convolutional} which has shown state of the art performance for sentiment classification on Twitter.
+The winning entry for SemEval15 \cite{severynunitn} task on Twitter sentiment classification also used DCNN.
+Figure \ref{fig:dcnn} summarizes the architecture.
+
+When used for sentiment classification on Twitter, the input to the DCNN is a matrix of word embeddings for each word in the tweet.
+For example, if the tweet consists of $s$ words then the input to the DCCNN is:
+\begin{flalign*}
+	S = 
+	\begin{bmatrix}
+	| & | & \ldots & | \\
+	w_1 & w_2 & \ldots & w_s \\
+	| & | & \ldots & | \\
+	\end{bmatrix}_{k\times s}
+\end{flalign*}
+where each $w_i \in R^k$ is a $k$-dimensional dense word embedding \cite{mikolov2013distributed}.
+The architecture consists of multiple layers of convolutions and max-pooling on top of the input matrix, followed a fully connected layer which is input to a softmax.
+The convolutions are of type \emph{wide-convolutions} of one-dimension.
+For example, for the input matrix $S \in R^{k\times s}$, a wide-convolution filter operating on $S$ will consist of convolution weights $m\in R^{k\times c}$ and will result in a matrix having dimension $k \times (s+c-1)$.
+Note here $c$ is the convolution filter width, which is a hyperparameter.
+The max-pooling operations presented in \cite{kalchbrenner2014convolutional} are different from the regular max-pooling.
+They present {\it dynamic $k$-max pooling}.
+$k$-max pooling takes the top $k$ maximum activations as opposed to just the maximum activation and the value of $k$ is selected dynamically based on the following formula: $k_l = \max \left( k_{top}, 	\lceil \frac{L-l}{L} s\rceil \right)$, where $l$ is number of current convolution layer, $L$ is total number of convolutions and $k_{top}$ is a fixed hyperparameter.
+Note that while \cite{kalchbrenner2014convolutional} used multiple layers of convolutions and max-pooling, subsequent work found that using a single layer of convolution and max-pooling gives similar results \cite{kim2014convolutional} \cite{severynunitn}.
+
+\begin{figure}
+	\centering
+	\includegraphics[height=4in, width=0.8\textwidth]{figs/DCNN.png}
+	\caption{Dynamic Convulational Neural Network of \cite{kalchbrenner2014convolutional} (Source: \cite{kalchbrenner2014convolutional})}
+	\label{fig:dcnn}
+\end{figure}
+
+\subsection{Recurrent Neural Networks with Long Short Term Memory (LSTM)}
+Recurrent Neural Networks (RNNs) are a class of artificial neural networks used for modeling sequences.
+RNNs are highly flexible in their use of context information as they can learn what part of the input sequence to store to memory and what parts to ignore.
+They also allow modeling of various regimes of sequence modeling as shown in Figure \ref{fig:rnn}.
+Please refer to \cite{graves2012supervised} for a comprehensive review of sequence modeling using RNN.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{figs/rnn.jpeg}
+	\caption{RNNs allow modeling of multiple types of input and output sequences (Source: \cite{karpathy-rnn})}
+	\label{fig:rnn}
+\end{figure}
+
+One of the short comings of RNN is that it is very difficult to store information over long sequences because of problems due to vanishing and exploding gradients as explained in \cite{hochreiter2001gradient}.
+{\it Long Short-Term Memory (LSTM)} \cite{hochreiter1997long} are designed to remedy this and store information over larger input sequences.
+They achieve this using special ``memory cell'' units. 
+Figure \ref{fig:lstm} shows the architecture of this cell which consists of an input gate, a forget gate, an output gate and a recurring cell state. 
+Refer to \cite{colah} for a gentle introduction to LSTM and to \cite{graves2012supervised} for a more comprehensive review and applications.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{figs/LSTM.png}
+	\caption{The repeating module of LSTM.
+$x_t$ is the input as time $t$ and $h_t$ is the output from the LSTM output gate at time $t$.
+The top horizontal line corresponds to the cell state and the bottom line corresponds to the hidden state (both of which are recurring states).
+	(Source: \cite{colah})}
+	\label{fig:lstm}
+\end{figure}
+
+\section{Twitter Sentiment Analysis using LSTM}
 

### 2015-12-20 paper/paper.tex
more details

@@
 
-For the latter, we compare initializing the word embeddings using random vectors (``word/rnd'') against initialization using the embeddings provided in \cite{tang2014learning} (``word/sswe'').
+We compare word-level models under two settings: initializing the word embeddings using random vectors (``word/rnd'') and initialization using the embeddings provided in \cite{tang2014learning} (``word/sswe'').
+
+We train all models using the implementation of the Adam algorithm \cite{kigma2014adam} provided in Lasagne \cite{lasagne} and a learning rate set to $0.1$ \footnote{We retrained the ascii/rnd/200 on SemEval using AdaGrad and a learning rate of $0.01$ to achieve $84.13$; using Adam and $0.1$ learning rate, the result was $83.21$}.
 
@@
 
+The character-level models using the ``ascii'' character set outperformed the other models.
+It is worth noting that the ``utf8'' model performaed comparably despite having a much larger vocabulary size.
+It is not suprising that the ``utf8'' model performed worse than the ``ascii'' model, as the test data consisted of only English tweets; however, since the ``utf8'' model is implicitly multi-lingual, our results suggest that the same model may perform well across multiple languages.
+We leave such experiments for future work.
+
 \subsection{Analysis}
+Figure ~\ref{fig:cool} shows the effect of character repetition on model confidence over the course of the sequence, where confidence is computed using the softmax function:
 
-\begin{figure}[h]
+$$P(y = j|\mathbf{x}) = \frac{e^{\mathbf{x}^T \mathbf{w}_j}}{\sum_{k=1}^{K} e^{\mathbf{x}^T \mathbf{w}_k}}$$
+
+In our experiments, we had $K = 2$ corresponding to binary classification between ``positive'' and ``negative'' tweets.
+Sequences ending in periods (``cool.'', ``coool.'') ended up with less-confident scores than tweets not ending in periods.
+Repeated exclamation points don't increase the model's confidence in the ``positive'' label as much as might be expected.
+
+\begin{figure}[h!]
+\label{fig:cool}
 \begin{center}
@@
 
-\begin{figure}[h]
+Figure ~\ref{fig:puppies} shows that the character-level model learns word meaning at a lexical level: the predictions for ``I love puppies'' and ``I hate puppies'' diverges sharply after the model has finished reading ``lov'' (``love'') and ``hat'' (``hate'').
+
+\begin{figure}[h!]
+\label{fig:puppies}
 \begin{center}
@@
 
-\begin{figure}[h]
-\begin{center}
-%\framebox[4.0in]{$\;$}
-\includegraphics[width=0.95\textwidth]{figs/cool}
-%\fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
-\end{center}
-\caption{Sample figure caption.}
-\end{figure}
+Figure ~\ref{fig:dentist} provides further evidence that character-level models can reason about lexical semantics in an intuitive fashion.
+We compare confidence contours across four tweets from the Go test set.
+The ground truth labels for the first two tweets are both ``negative'', and for the second two are both ``positive''.
+In the first two, the model finds strong evidence for a ``negative'' prediction before reaching the word ``dentist'', and correctly predicts that both tweets are ``negative''.
+In the second two, the word ``dentist'' results in an increase in the model's confidence in a ``negative'' prediction; however, the words ``enjoyable'' and ``:)'' cause the model's confidence to decrease in the ``positive'' direction.
 
-\begin{figure}[h]
+\begin{figure}[h!]
+\label{fig:dentist}
 \begin{center}

### 2015-12-20 paper/paper.tex
update: conclusion, merge in Jun paper

@@
 \usepackage{url}
+\usepackage{amsmath}
 %\documentstyle[nips14submit_09,times,art10]{article} % For LaTeX 2.09
@@
 \section{Introduction}
+% why analysis sentiment expecially twitter
+In the last few years, microblogging has become a very popular communication tool in people's social life.
+On microblogging platforms, such as Twitter\footnote{\url{https://twitter.com/}} and Facebook\footnote{\url{https://www.facebook.com/}}, a diverse range of people are attracted to post short sentences, images, and video links to share life issues and opinions.
+This popularity results in enormous amount of information covering a wide range of topics on the brands, products, politics and social events. 
+Such data is a valuable and efficient source for marketing and social studies. 
+Sentiment analysis on microblogging has obtained special interest, because it determines the attitude of a user with respect to some topic or product and thus provides provide convincing information
+For example, it may help manufacturing companies to know how people like their product (or service), what would people prefer.
 
-We use LSTM \cite{hochreiter1997long} to do stuff
+In this paper, we focus on using sentiment analysis on Twitter. 
+Twitter is an extremely popular microblogging platform which allows people to post messages of up to 140 characters.
+Because of the short nature of tweets, people often post twitter messages (called Tweets) frequently while attending events like product launches, movie premiers, music concerts or just to express their opinion on a trending or current topic.
+As such, they can be a valuable source of public opinion or feedback \cite{o2010tweets, bollen2011twitter, bollen2009modeling}.
+
+Realizing this importance, analyzing sentiment of Tweets has been a recurring task in the SemEval \footnote{\url{http://alt.qcri.org/semeval2016/task4/}} competitions.
+
+%We restrict our interest on Twitter for the following reasons: 
+%People are posting on various topics on Twitter, thus it is a valuable source of data; Twitter is one of the most popular microblogging platform and the number of messages are growing everyday.
+The collected corpus can be arbitrary large.
+Users of Twitter have various background and have different language conventions.
+It is possible to collect tweets of wide diversity in both topics and language models.
+
+Although there exist plenty of work on text classification, some unique characteristics of tweets present special challenges for sentiment analysis:
+1.
+Tweets are short in length.
+There is a limitation of $140$ words for each tweet; 2.
+The language used in tweets is very informal with misspelling, creative spelling,  new words, slangs, and URLs; 3.
+Emotions and hashtags are frequently used.
+
+% what we did
+In this paper, we propose a bi-directional Long Short-Term Memory (LSTM) method for sentiment analysis.
+We tried both word-level and character-level features on the bi-directional LSTM model and compare the results with Dynamic Convolutional Neural Network (DCNN) \cite{kalchbrenner2014convolutional}.
+We show that the accuracy of sentiment analysis ... (Please revise this part.)
+
+We train our model on $1.6$ M distantly-supervised tweets collected by Go et.
+al.
+\cite{go2009twitter}, and evaluate the results on SemEval-2016 Task 4\footnote{\url{http://alt.qcri.org/semeval2016/task4/}}.
+Currently we focus on polarity classification, and plan to apply our model on 5-point scale classification in the future.
+
+% the structure of the paper
+In the remaining of this paper, we first introduce the related work, the dataset, and the evaluation methodology.
+We then describe the model we proposed.
+Finally, we discuss the experiment results and point out possible directions for future research.
+(Please revise this part.)
 
 \section{Related Work}
+Twitter sentiment analysis is increasingly drawing attention of researchers in recent years. 
+Given the length limitations on tweets, sentiment analysis of tweets is often considered similar to sentence-level sentiment analysis \cite{kouloumpis2011twitter}.
+However, phrase and sentence level approaches can hardly define the sentiment of some specific topics.
+Considering opinions adhering on different topics, Wang et.
+al.
+\cite{wang2011topic} proposed a hashtag-level sentiment classification method  to generate the overall sentiment polarity for a given hashtag.
+Recently, following the work of \cite{mikolov2013efficient} some researchers used neural network to implement sentiment classification. 
+For example, Kim \cite{kim2014convolutional} adopted convolutional neural networks to learn sentiment-bearing sentence vectors, Mikolov et al.
+\cite{mikolov2013distributed} proposed Paragraph vector which outperformed bag-of-words model for sentiment analysis, and Tang et.
+al.
+\cite{tang2014learning} used ConvNets to learn sentiment specific word embedding (SSWE), which encodes sentiment information in the continuous
+representation of words.
+Furthermore, Kalchbrenner \cite{kalchbrenner2014convolutional} proposed a Dynamic Convolutional Neural Network (DCNN) which uses dynamic k-max pooling, a global pooling operation over linear sequences.
+Instead of directly applying ConvNets to embeddings of words, \cite{zhang2015character} applies the network only on characters.
+They showed that the deep ConvNets does not require knowledge of words and thus can work for different languages.
+LSTM \cite{hochreiter1997long} is another state-of-the-art semantic composition models for sentiment classification \cite{li2015tree}.
+Similar to DCNN, it also learns fixed-length vectors for sentences of varying length, captures words order in a sentence and does not depend on external dependency or constituency parse results.
 
-Dynamic convolutional neural net for sentences \cite{kalchbrenner2014convolutional}
+\subsection{Dynamic Convolutional Neural Networks (DCNN)}
+We briefly review the architecture of DCNN \cite{kalchbrenner2014convolutional} which has shown state of the art performance for sentiment classification on Twitter.
+The winning entry for SemEval15 \cite{severynunitn} task on Twitter sentiment classification also used DCNN.
+Figure \ref{fig:dcnn} summarizes the architecture.
 
-sentiment specific word embeddings \cite{tang2014learning}
+When used for sentiment classification on Twitter, the input to the DCNN is a matrix of word embeddings for each word in the tweet.
+For example, if the tweet consists of $s$ words then the input to the DCCNN is:
+\begin{flalign*}
+	S = 
+	\begin{bmatrix}
+	| & | & \ldots & | \\
+	w_1 & w_2 & \ldots & w_s \\
+	| & | & \ldots & | \\
+	\end{bmatrix}_{k\times s}
+\end{flalign*}
+where each $w_i \in R^k$ is a $k$-dimensional dense word embedding \cite{mikolov2013distributed}.
+The architecture consists of multiple layers of convolutions and max-pooling on top of the input matrix, followed a fully connected layer which is input to a softmax.
+The convolutions are of type \emph{wide-convolutions} of one-dimension.
+For example, for the input matrix $S \in R^{k\times s}$, a wide-convolution filter operating on $S$ will consist of convolution weights $m\in R^{k\times c}$ and will result in a matrix having dimension $k \times (s+c-1)$.
+Note here $c$ is the convolution filter width, which is a hyperparameter.
+The max-pooling operations presented in \cite{kalchbrenner2014convolutional} are different from the regular max-pooling.
+They present {\it dynamic $k$-max pooling}.
+$k$-max pooling takes the top $k$ maximum activations as opposed to just the maximum activation and the value of $k$ is selected dynamically based on the following formula: $k_l = \max \left( k_{top}, 	\lceil \frac{L-l}{L} s\rceil \right)$, where $l$ is number of current convolution layer, $L$ is total number of convolutions and $k_{top}$ is a fixed hyperparameter.
+Note that while \cite{kalchbrenner2014convolutional} used multiple layers of convolutions and max-pooling, subsequent work found that using a single layer of convolution and max-pooling gives similar results \cite{kim2014convolutional} \cite{severynunitn}.
 
-character-lvl CNN for text classification \cite{zhang2015character}
+\begin{figure}
+	\centering
+	\includegraphics[height=4in, width=0.8\textwidth]{figs/DCNN.png}
+	\caption{Dynamic Convulational Neural Network of \cite{kalchbrenner2014convolutional} (Source: \cite{kalchbrenner2014convolutional})}
+	\label{fig:dcnn}
+\end{figure}
 
-\section{Model}
+\subsection{Recurrent Neural Networks with Long Short Term Memory (LSTM)}
+Recurrent Neural Networks (RNNs) are a class of artificial neural networks used for modeling sequences.
+RNNs are highly flexible in their use of context information as they can learn what part of the input sequence to store to memory and what parts to ignore.
+They also allow modeling of various regimes of sequence modeling as shown in Figure \ref{fig:rnn}.
+Please refer to \cite{graves2012supervised} for a comprehensive review of sequence modeling using RNN.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{figs/rnn.jpeg}
+	\caption{RNNs allow modeling of multiple types of input and output sequences (Source: \cite{karpathy-rnn})}
+	\label{fig:rnn}
+\end{figure}
+
+One of the short comings of RNN is that it is very difficult to store information over long sequences because of problems due to vanishing and exploding gradients as explained in \cite{hochreiter2001gradient}.
+{\it Long Short-Term Memory (LSTM)} \cite{hochreiter1997long} are designed to remedy this and store information over larger input sequences.
+They achieve this using special ``memory cell'' units. 
+Figure \ref{fig:lstm} shows the architecture of this cell which consists of an input gate, a forget gate, an output gate and a recurring cell state. 
+Refer to \cite{colah} for a gentle introduction to LSTM and to \cite{graves2012supervised} for a more comprehensive review and applications.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{figs/LSTM.png}
+	\caption{The repeating module of LSTM.
+$x_t$ is the input as time $t$ and $h_t$ is the output from the LSTM output gate at time $t$.
+The top horizontal line corresponds to the cell state and the bottom line corresponds to the hidden state (both of which are recurring states).
+	(Source: \cite{colah})}
+	\label{fig:lstm}
+\end{figure}
 
@@
 
-We compare word-level models under two settings: initializing the word embeddings using random vectors (``word/rnd'') and initialization using the embeddings provided in \cite{tang2014learning} (``word/sswe'').
+We compare word-level models under two settings: initializing the word embeddings using random vectors (``word/rnd'') and initialization using ``sentiment-specific'' word embeddings provided in \cite{tang2014learning} (``word/sswe'').
 
-We train all models using the implementation of the Adam algorithm \cite{kigma2014adam} provided in Lasagne \cite{lasagne} and a learning rate set to $0.1$ \footnote{We retrained the ascii/rnd/200 on SemEval using AdaGrad and a learning rate of $0.01$ to achieve $84.13$; using Adam and $0.1$ learning rate, the result was $83.21$}.
+Our results are shown in Table ~\ref{table:results}.
+We train all models using the implementation of the Adam algorithm \cite{kingma2014adam} provided in Lasagne\footnote{\url{https://github.com/Lasagne/Lasagne}} and a learning rate set to $0.1$ \footnote{We retrained the ascii/rnd/200 on SemEval using AdaGrad and a learning rate of $0.01$ to achieve $84.13$; using Adam and $0.1$ learning rate, the result was $83.21$}.
+To learn word and character embeddings, we use a bidirectional LSTM with $256$ hidden units, followed by a mean pooling and a dropout layer ($p=0.5$), a second forward-directional LSTM (again with $256$ hidden units) and a final dropout layer ($p=0.6$).
+We obtain final predictions using the softmax function.
 
 \begin{table}[h!]
+\label{table:results}
 \centering
 \caption{Accuracy across LSTMs}
-\label{char-models}
 \begin{tabular}{|l|ll|}
@@
 \hline
-ascii/rnd/50  & \textbf{83.84} & 82.08   \\
+ascii/rnd/50  & 83.84          & 82.08            \\
 ascii/rnd/200 & 82.45          & \textbf{84.13}   \\
@@
 utf8          & 81.34          & 82.34   \\
+char-dcnn     & 75.0           & 81.3    \\
 word/rnd      & 81.85          & 78.07   \\
 word/sswe     & 83.24          & 79.27   \\
+word/dcnn     & \textbf{87.4}\footnote{Reported in \cite{kalchbrenner2014convolutional}}           & 81.3    \\
 \hline
@@
 
-The character-level models using the ``ascii'' character set outperformed the other models.
-It is worth noting that the ``utf8'' model performaed comparably despite having a much larger vocabulary size.
+The character-level models using the ``ascii'' character set outperformed the other models on the SemEval dataset.
+Due to the highly-productive nature of the Twitter ``lexicon'', users' predisposition toward using slang dialects, and the constraint of the $140$-character limit, it makes sense that word-level models underperform.
+
+It is worth noting that the ``utf8'' model performed comparably despite having a much larger vocabulary size.
 It is not suprising that the ``utf8'' model performed worse than the ``ascii'' model, as the test data consisted of only English tweets; however, since the ``utf8'' model is implicitly multi-lingual, our results suggest that the same model may perform well across multiple languages.
@@
 
-\subsection{Analysis}
+As Table ~\ref{table:results} shows, \cite{kalchbrenner2014convolutional} outperform our character LSTM on the Go dataset.
+However, the results presented for our character and word-level models achieve competitive performance on the Go dataset without tuning hyperparameters such as learning rate and network width.
+
+\subsection{Qualitative Analysis}
 Figure ~\ref{fig:cool} shows the effect of character repetition on model confidence over the course of the sequence, where confidence is computed using the softmax function:
@@
 \end{center}
-\caption{Sample figure caption.}
+\caption{Comparison of model confidence for different forms of the word ``cool''}
 \end{figure}
@@
 \end{center}
-\caption{Sample figure caption.}
+\caption{Comparison of model confidence for ``I love puppies'' vs. ``I hate puppies''}
 \end{figure}
@@
 \end{center}
-\caption{Confidence in ``negative'' prediction over time}
+\caption{Confidence in ``negative'' prediction over time across four tweets}
 \end{figure}
@@
 \section{Conclusion}
-
-\subsubsection*{Acknowledgments}
+In this paper, we show promising empirical evidence that character-level models perform well for sentiment analysis on tweets, which are ``noisy'', in that users think of millions of new words and spellings of words every day, and ``brief'', in that they are constrained to $140$ characters.
+This type of text data is becoming an important research focus in the field of NLP, as data is often cheap to collect in high volume and can provide important insight into society-wide trends.
+Though the experiments in this paper focus on sentiment classification, we believe our results provide a basis for future work on character-level modelling for a variety of other NLP tasks.
 

### 2015-12-20 paper/paper.tex
change some fig labels

@@
 \begin{abstract}
-abstract goes here
+We show promising empirical evidence that character-level LSTMs perform well for sentiment analysis on tweets, which are ``noisy'', in that users think of millions of new words and spellings of words every day, and ``brief'', in that they are constrained to $140$ characters.
+This type of text data is becoming an important research focus in the field of NLP, as data is often cheap to collect in high volume and can provide important insight into society-wide trends.
+Our model achieves $84 \%$ accuracy, a result competitive with the state of the art for binary sentiment classification on tweets.
 \end{abstract}
@@
 \caption{Data size and label distribution}
-\label{my-label}
 \begin{tabular}{|lllll|}
@@
 \end{tabular}
+\label{table:corpus}
 \end{table}
@@
 \begin{table}[h!]
-\label{table:results}
 \centering
@@
 \end{tabular}
+\label{table:results}
 \end{table}
@@
 \begin{figure}[h!]
-\label{fig:cool}
 \begin{center}
@@
 \caption{Comparison of model confidence for different forms of the word ``cool''}
+\label{fig:cool}
 \end{figure}
@@
 \begin{figure}[h!]
-\label{fig:puppies}
+
 \begin{center}
@@
 \caption{Comparison of model confidence for ``I love puppies'' vs. ``I hate puppies''}
+\label{fig:puppies}
 \end{figure}
@@
 \begin{figure}[h!]
-\label{fig:dentist}
+
 \begin{center}
@@
 \caption{Confidence in ``negative'' prediction over time across four tweets}
+\label{fig:dentist}
 \end{figure}
@@
 This type of text data is becoming an important research focus in the field of NLP, as data is often cheap to collect in high volume and can provide important insight into society-wide trends.
-Though the experiments in this paper focus on sentiment classification, we believe our results provide a basis for future work on character-level modelling for a variety of other NLP tasks.
-
-\subsubsection*{References}
+Though the experiments in this paper focus on sentiment classification, we believe our results provide a basis for future work on character-level modeling for a variety of other NLP tasks.
 

### 2015-12-20 paper-Jun/paper.tex
model section

@@
 \section{Twitter Sentiment Analysis using LSTM}
+In this section we present LSTM models for sentiment analysis of twitter messages.
+We explored different architectures of LSTM networks which operate at word-level or character-level input.
+The basic architecture of the network is shown in Figure \ref{fig:mylstm}.
+We explain the model architecture considering the character input.
+Given a tweet as a sequence of characters $X=\{x_1, \ldots, x_{140}\}$, the task is to predict the sentiment of Tweet as being {\it positive} (1) or {\it negative} (0).
+Each $x_i$ is a one-hot encoding of either the character or the word.
+The LSTM models then take this one-hot encoding and convert them into either a {\it character embedding} or a {\it word embedding} depending on whether the input is characters or words.
+The embeddings can be randomly initialized and learned jointly with other model parameters.
+
+The model consists of a bidirectional LSTM layer over this character input $X$.
+The hidden layer activations obtained from the forward and backward LSTM layer are averaged at {\it each} character which serves as the input to the {\it second layer} of LSTM.
+The second layer of LSTM takes as input this sequence of hidden layer outputs from first layer and encodes a representation of the tweet at the last element of the sequence (the hidden layer activation of the last LSTM unit).
+This output is then fed into a softmax layer which classifies the Tweet as being positive or negative.
+The model is trained using categorical cross-entropy.
+
+Note that it is possible to have multiple levels of granularity in predictions, like positive, ``netural" and negative, or even finer.
+We restrict ourselves to two classes to be able to compare with the state of the art published results \cite{kalchbrenner2014convolutional}.
+
+The same model can be used with either character input data or word input data.
+Note that when used with word input, it is common to initialize the models with pre-trained word embeddings \cite{mikolov2013distributed}.
+We explore multiple such initializations in our experiments.
+For the character input model, the embeddings are always initialized randomly, since pre-trained character embeddings are not yet available.
+
+We explored other variants of the model like have single-directional LSTM, having a single layer as opposed to multiple layers, using concatenation instead of averaging after first layer.
+We also tried an ensemble LSTM which takes both characters and words as separate inputs which are combined at the last stage by concatenation before feeding to the softmax.
+Unfortunately, none of these models gave good preliminary results and in the experiments we focused on just the model of Figure \ref{fig:mylstm} with either character or words as inputs.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth, height=3.6in]{figs/mylstm.pdf}
+	\caption{The bi-directional LSTM model used for Twitter sentiment analysis. Each circular node represents an LSTM cell. The input to the model is the sequence $x$. }
+	\label{fig:mylstm}
+\end{figure}
 
@@
 \end{table}
-
-This is a figure:
-
-\begin{figure}[h]
-\begin{center}
-%\framebox[4.0in]{$\;$}
-\fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
-\end{center}
-\caption{Sample figure caption.}
-\end{figure}
 

### 2015-12-20 paper/paper.tex
add similar words table

@@
 
+\section{Twitter Sentiment Analysis using LSTM}
+In this section we present LSTM models for sentiment analysis of twitter messages.
+We explored different architectures of LSTM networks which operate at word-level or character-level input.
+The basic architecture of the network is shown in Figure \ref{fig:mylstm}.
+We explain the model architecture considering the character input.
+Given a tweet as a sequence of characters $X=\{x_1, \ldots, x_{140}\}$, the task is to predict the sentiment of Tweet as being {\it positive} (1) or {\it negative} (0).
+Each $x_i$ is a one-hot encoding of either the character or the word.
+The LSTM models then take this one-hot encoding and convert them into either a {\it character embedding} or a {\it word embedding} depending on whether the input is characters or words.
+The embeddings can be randomly initialized and learned jointly with other model parameters.
+
+The model consists of a bidirectional LSTM layer over this character input $X$.
+The hidden layer activations obtained from the forward and backward LSTM layer are averaged at {\it each} character which serves as the input to the {\it second layer} of LSTM.
+The second layer of LSTM takes as input this sequence of hidden layer outputs from first layer and encodes a representation of the tweet at the last element of the sequence (the hidden layer activation of the last LSTM unit).
+This output is then fed into a softmax layer which classifies the Tweet as being positive or negative.
+The model is trained using categorical cross-entropy.
+
+Note that it is possible to have multiple levels of granularity in predictions, like positive, ``netural" and negative, or even finer.
+We restrict ourselves to two classes to be able to compare with the state of the art published results \cite{kalchbrenner2014convolutional}.
+
+The same model can be used with either character input data or word input data.
+Note that when used with word input, it is common to initialize the models with pre-trained word embeddings \cite{mikolov2013distributed}.
+We explore multiple such initializations in our experiments.
+For the character input model, the embeddings are always initialized randomly, since pre-trained character embeddings are not yet available.
+
+We explored other variants of the model like have single-directional LSTM, having a single layer as opposed to multiple layers, using concatenation instead of averaging after first layer.
+We also tried an ensemble LSTM which takes both characters and words as separate inputs which are combined at the last stage by concatenation before feeding to the softmax.
+Unfortunately, none of these models gave good preliminary results and in the experiments we focused on just the model of Figure \ref{fig:mylstm} with either character or words as inputs.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth, height=3.6in]{figs/mylstm.pdf}
+	\caption{The bi-directional LSTM model used for Twitter sentiment analysis. Each circular node represents an LSTM cell. The input to the model is the sequence $x$. }
+	\label{fig:mylstm}
+\end{figure}
+
 \section{Results}
@@
 
+Table ~\ref{table:word-sim} shows words alongside their nearest neighbors computed using the character embeddings learned by our model.
+As desired, othographic variations of the same word ``type'' are close together in the embedding space.
+
+\begin{table}[h!]
+\begin{center}
+\begin{tabular}{|l|l|}
+\hline
+woohoo &    wooohoo woohooo wooohooo woohoooo wooooo whoooo woooo whoohoo \\
+wohooo
+woww &   wowww wow whack vwv waw vry cow werk woke \\
+xoxo &    xox xoxox zoo xxoo vox xoxoxo zoe xxo twix \\
+thanks &    thankss thanku yanks thankx hanks thankz thanksss tanks thanx \\
+sorry &    sorrry srry scary sowwy sorryyy scared sory scarred errrr \\
+sucks &   sucky suckss yucky eurgh ouchy suckish fucks ouchh jerks \\
+twitter &    twiter twiiter tinytwitter twitterr quicker txting \\
+granddaughter & ktbeeper twitting \\
+tweet  &  tweeet tweep tweety tweets textt theee seee outer peace \\
+\hline
+\end{tabular}
+\end{center}
+\caption{Words and their nearest neighbors}
+\label{table:word-sim}
+\end{table}
+
 \section{Conclusion}

### 2015-12-20 paper-Jun/paper.tex
intro

@@
 
-Although there exist plenty of work on text classification, some unique characteristics of tweets present special challenges for sentiment analysis:
-1.
-Tweets are short in length.
-There is a limitation of $140$ words for each tweet; 2.
-The language used in tweets is very informal with misspelling, creative spelling,  new words, slangs, and URLs; 3.
-Emotions and hashtags are frequently used.
+Although there exists plenty of work on text classification, some unique characteristics of tweets present special challenges for sentiment analysis:
+1) Tweets are short in length.
+There is a limitation of $140$ words for each tweet which makes analyzing them challenging \cite{mehrotra2013improving}; 2) The language used in tweets is very informal with misspellings (often intentional, like different spellings of "cool": coool, coolll, coooolll!!), new words, slangs, and URLs;
+3) The number of tweets increases at a very fast pace, and with new data comes new words, new trends in using abbreviations, which lead to a frequent problem of out-of-vocabulary words
+4) Special symbols and their combinations are often used, like emoticons and hashtags.
 
 % what we did
-In this paper, we propose a bi-directional Long Short-Term Memory (LSTM) method for sentiment analysis.
-We tried both word-level and character-level features on the bi-directional LSTM model and compare the results with Dynamic Convolutional Neural Network (DCNN) \cite{kalchbrenner2014convolutional}.
-We show that the accuracy of sentiment analysis ... (Please revise this part.)
+Keeping in view the challenges posed by the nature of tweets, 
+in this paper, we propose to use bi-directional Long Short-Term Memory (LSTM) 
+networks which operate at {\it character-level input} and make predictions at the tweet-level.
+Such networks naturally handle the problem of very large vocabulary sizes and the presence of sub-word information, without having to keep many trained embedding which would result from keeping very large vocabularies.
+We find that the model interestingly performs {\bf better} than equivalent LSTM model which operates at the word level (even when initialized with a variety of pre-trained embeddings).
+We also compare our results with Dynamic Convolutional Neural Network (DCNN) \cite{kalchbrenner2014convolutional} which has shown state of the art performance on Twitter sentiment classification \cite{severynunitn}.
+We explore the properties of the character LSTM model and find that it learns to find meaning in composition of characters and can effectively relate sub-word information (like presence of multiple exclamation marks) to the sentiment of the tweet.
 
-We train our model on $1.6$ M distantly-supervised tweets collected by Go et.
+%We train our model on $1.6$ M distantly-supervised tweets collected by Go et.
 al.
-\cite{go2009twitter}, and evaluate the results on SemEval-2016 Task 4\footnote{\url{http://alt.qcri.org/semeval2016/task4/}}.
+\cite{go2009twitter}, and evaluate the results on SemEval-2016 Task 4.
 Currently we focus on polarity classification, and plan to apply our model on 5-point scale classification in the future.
@@
 
-\section{Twitter Sentiment Analysis using LSTM}
+\section{Twitter Sentiment Analysis using Character LSTM}
 In this section we present LSTM models for sentiment analysis of twitter messages.

### 2015-12-20 paper/paper.tex
update

@@
 As such, they can be a valuable source of public opinion or feedback \cite{o2010tweets, bollen2011twitter, bollen2009modeling}.
-
 Realizing this importance, analyzing sentiment of Tweets has been a recurring task in the SemEval \footnote{\url{http://alt.qcri.org/semeval2016/task4/}} competitions.
@@
 % the structure of the paper
-In the remaining of this paper, we first introduce the related work, the dataset, and the evaluation methodology.
-We then describe the model we proposed.
-Finally, we discuss the experiment results and point out possible directions for future research.
-(Please revise this part.)
 
@@
 woohoo &    wooohoo woohooo wooohooo woohoooo wooooo whoooo woooo whoohoo \\
-wohooo
+wohooo\
 woww &   wowww wow whack vwv waw vry cow werk woke \\

### 2015-12-20 paper/paper.tex
add the 'in this paper' part of the intro

@@
 % the structure of the paper
+This paper is laid out as follows: in Section 2, we describe prior work on sentiment classification using deep learning.
+In Section 3, we describe the structure of our model.
+In Section 4, we describe the datasets used to train and evaluate our model, as well as results from several experiments.
+We conclude with an analysis of our results, as well as describe areas of future research.
 

### 2015-12-20 paper/paper.tex

@@
 
-\title{LSTM for Sentiment Analysis on Twitter}
+\usepackage{cleveref}
+\usepackage{xcolor}
+\hypersetup{
+	colorlinks,
+	linkcolor={red!50!black},
+	citecolor={blue!50!black},
+	urlcolor={blue!80!black}
+}
+
+\title{Character LSTM for Sentiment Analysis on Twitter}
 
@@
 In the last few years, microblogging has become a very popular communication tool in people's social life.
-On microblogging platforms, such as Twitter\footnote{\url{https://twitter.com/}} and Facebook\footnote{\url{https://www.facebook.com/}}, a diverse range of people are attracted to post short sentences, images, and video links to share life issues and opinions.
-This popularity results in enormous amount of information covering a wide range of topics on the brands, products, politics and social events. 
+On microblogging platforms, such as Twitter\footnote{\url{https://twitter.com/}} and Facebook\footnote{\url{https://www.facebook.com/}}, a diverse range of people are attracted to post short sentences, images, and video links to comment on and share their opinion on a variety of topics.
+This popularity results in enormous amount of information covering a wide range of topics on various brands, products, politics and social events. 
 Such data is a valuable and efficient source for marketing and social studies. 
-Sentiment analysis on microblogging has obtained special interest, because it determines the attitude of a user with respect to some topic or product and thus provides provide convincing information
-For example, it may help manufacturing companies to know how people like their product (or service), what would people prefer.
+Sentiment analysis on microblogging has obtained special interest, because it determines the attitude of a user with respect to some topic or product and thus provides convincing information of user's interest or feedback on the topic.
+For example, it may help a consumer products company get relevant feedback of public opinion on a recent product launch.
 
-In this paper, we focus on using sentiment analysis on Twitter. 
-Twitter is an extremely popular microblogging platform which allows people to post messages of up to 140 characters.
+In this paper, we focus on sentiment analysis on Twitter. 
+Twitter is an extremely popular microblogging platform which allows users to post messages of up to 140 characters.
 Because of the short nature of tweets, people often post twitter messages (called Tweets) frequently while attending events like product launches, movie premiers, music concerts or just to express their opinion on a trending or current topic.
-As such, they can be a valuable source of public opinion or feedback \cite{o2010tweets, bollen2011twitter, bollen2009modeling}.
+As such, they can be a valuable source of public opinion or feedback and have received considerable research interest, exemplary work among which are \cite{o2010tweets, bollen2011twitter, bollen2009modeling}.
 Realizing this importance, analyzing sentiment of Tweets has been a recurring task in the SemEval \footnote{\url{http://alt.qcri.org/semeval2016/task4/}} competitions.
@@
 Keeping in view the challenges posed by the nature of tweets, 
-in this paper, we propose to use bi-directional Long Short-Term Memory (LSTM) 
-networks which operate at {\it character-level input} and make predictions at the tweet-level.
-Such networks naturally handle the problem of very large vocabulary sizes and the presence of sub-word information, without having to keep many trained embedding which would result from keeping very large vocabularies.
-We find that the model interestingly performs {\bf better} than equivalent LSTM model which operates at the word level (even when initialized with a variety of pre-trained embeddings).
+in this paper, we propose to use bi-directional Long Short-Term Memory (LSTM) \cite{hochreiter1997long} 
+networks which operate at {\it character-level input} and make predictions at the Tweet-level.
+Such networks naturally handle the problem of very large vocabulary sizes and the presence of sub-word information, without having to keep many trained embedding which would be required for word-level models  with very large vocabularies.
+We find that the character-level model, interestingly, performs {\bf better} than equivalent LSTM model which operates at the word level (even when initialized with a variety of pre-trained embeddings).
 We also compare our results with Dynamic Convolutional Neural Network (DCNN) \cite{kalchbrenner2014convolutional} which has shown state of the art performance on Twitter sentiment classification \cite{severynunitn}.
-We explore the properties of the character LSTM model and find that it learns to find meaning in composition of characters and can effectively relate sub-word information (like presence of multiple exclamation marks) to the sentiment of the tweet.
+In addition, we also explore the properties of the character LSTM model and find that it learns to find meaning in composition of characters and can effectively relate sub-word information (like presence of period instead of exclamation marks) to the sentiment of the tweet.
 
@@
 Furthermore, Kalchbrenner \cite{kalchbrenner2014convolutional} proposed a Dynamic Convolutional Neural Network (DCNN) which uses dynamic k-max pooling, a global pooling operation over linear sequences.
-Instead of directly applying ConvNets to embeddings of words, \cite{zhang2015character} applies the network only on characters.
+Instead of directly applying ConvNets to embeddings of words, very recently \cite{zhang2015character} applied the network only on characters.
 They showed that the deep ConvNets does not require knowledge of words and thus can work for different languages.
 LSTM \cite{hochreiter1997long} is another state-of-the-art semantic composition models for sentiment classification \cite{li2015tree}.
-Similar to DCNN, it also learns fixed-length vectors for sentences of varying length, captures words order in a sentence and does not depend on external dependency or constituency parse results.
+Similar to DCNN, it also learns fixed-length vectors for sentences of varying length of word-level input, captures words order in a sentence and does not depend on external dependency or constituency parse results.
+
+Recently, there has been a surge of research interest in exploring character-input models for a variety of natural language processing tasks like: language modeling \cite{kim2015character}, POS tagging \cite{ling2015finding}, dependency parsing \cite{ballesteros2015improved} and machine translation \cite{ling2015character}.
+Our exploration is in a similar direction, focused on microblogging data which is naturally morphologically rich, and on using LSTM which are powerful models of sequence data and hence a natural choice for text modeling.
+
+Before explaining the architecture of our character LSTM model, we briefly describe the DCNN architecture and review Long Short Term Memory Networks.
 
@@
 We restrict ourselves to two classes to be able to compare with the state of the art published results \cite{kalchbrenner2014convolutional}.
+
+Intuitively, the model operating at character level is expected to get information of the characters of a word from the bidirectional character LSTMs which are then composed into a word-level meaning by the second layer of the LSTM.
+This LSTM layer then encodes the information of the entire tweet in a $r$ (=256 in experiments) dimensional space which can be used to classify the tweet.
 
@@
 
-Table ~\ref{table:word-sim} shows words alongside their nearest neighbors computed using the character embeddings learned by our model.
-As desired, othographic variations of the same word ``type'' are close together in the embedding space.
+To further analyze whether related spellings or variants of a word are mapped to nearby points in the final encoding, we fed as input individual words (from the dataset vocabulary) to the character input LSTM model and obtained the encoding of the word after the last layer (before the softmax).
+We then find the closest neighbors of every word in the encoding space.
+Table ~\ref{table:word-sim} shows words alongside their top nearest neighbors where closeness is measured by cosine similarity.
+As desired, orthographic variations of the same word ``type'' are close together in the embedding space.
+In particular, it is worth noting that "woohoo" with different number of ``o'' are all close.
+Similarly, the word "sowwy" which is a common way of saying ``sorry" in such informal conversations is also close to the correct word.
 
@@
 \hline
-woohoo &    wooohoo woohooo wooohooo woohoooo wooooo whoooo woooo whoohoo \\
-wohooo\
-woww &   wowww wow whack vwv waw vry cow werk woke \\
+Word & Nearest Neighbors \\ \hline
+woohoo &    wooohoo woohooo wooohooo woohoooo wooooo whoooo woooo whoohoo wohooo\\
+%woww &   wowww wow whack vwv waw vry cow werk woke \\
 xoxo &    xox xoxox zoo xxoo vox xoxoxo zoe xxo twix \\
@@
 sucks &   sucky suckss yucky eurgh ouchy suckish fucks ouchh jerks \\
-twitter &    twiter twiiter tinytwitter twitterr quicker txting \\
-granddaughter & ktbeeper twitting \\
+%twitter &    twiter twiiter tinytwitter twitterr quicker txting \\
+%granddaughter & ktbeeper twitting \\
 tweet  &  tweeet tweep tweety tweets textt theee seee outer peace \\

### 2015-12-20 paper/paper.tex
final sub
add another fig

@@
 
-%% \author{
-%% David S.~Hippocampus\thanks{ Use footnote for providing further information
-%% about author (webpage, alternative address)---\emph{not} for acknowledging
-%% funding agencies.} \\
-%% Department of Computer Science\\
-%% Cranberry-Lemon University\\
-%% Pittsburgh, PA 15213 \\
-%% \texttt{hippo@cs.cranberry-lemon.edu} \\
-%% \And
-%% Coauthor \\
-%% Affiliation \\
-%% Address \\
-%% \texttt{email} \\
-%% \AND
-%% Coauthor \\
-%% Affiliation \\
-%% Address \\
-%% \texttt{email} \\
-%% \And
-%% Coauthor \\
-%% Affiliation \\
-%% Address \\
-%% \texttt{email} \\
-%% \And
-%% Coauthor \\
-%% Affiliation \\
-%% Address \\
-%% \texttt{email} \\
-%% (if needed)\\
-%% }
-
 % The \author macro works with any number of authors. There are two commands
@@
 In the last few years, microblogging has become a very popular communication tool in people's social life.
-On microblogging platforms, such as Twitter\footnote{\url{https://twitter.com/}} and Facebook\footnote{\url{https://www.facebook.com/}}, a diverse range of people are attracted to post short sentences, images, and video links to comment on and share their opinion on a variety of topics.
+Microblogging platforms like Twitter and Facebook allow users to share their opinions on a wide variety of topics.
 This popularity results in enormous amount of information covering a wide range of topics on various brands, products, politics and social events.
-Such data is a valuable and efficient source for marketing and social studies. 
+
+Sentiment analysis on microblogging corpora has become an import area of research across the fields of NLP, machine learning, and social science research. 
 Sentiment analysis on microblogging has obtained special interest, because it determines the attitude of a user with respect to some topic or product and thus provides convincing information of user's interest or feedback on the topic.
+
 For example, it may help a consumer products company get relevant feedback of public opinion on a recent product launch.
@@
 \begin{figure}[h!]
-
 \begin{center}
@@
 \label{fig:dentist}
+\end{figure}
+
+Figure ~\ref{fig:basketball} shows that our model learns the effect of negation.
+
+\begin{figure}[h!]
+\begin{center}
+%\framebox[4.0in]{$\;$}
+\includegraphics[width=0.95\textwidth]{figs/basketball}
+%\fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
+\end{center}
+\caption{Confidence in ``positive'' prediction over time across four tweets}
+\label{fig:basketball}
 \end{figure}

### 2015-12-20 draft2/paper.tex

--- 2015-12-20 paper/paper.tex
@@
 \newcommand{\new}{\marginpar{NEW}}
+\newcommand{\todo}[1]{\textcolor{red}{TODO: #1}}
 
@@
 % why analysis sentiment expecially twitter
-In the last few years, microblogging has become a very popular communication tool in people's social life.
-Microblogging platforms like Twitter and Facebook allow users to share their opinions on a wide variety of topics.
-This popularity results in enormous amount of information covering a wide range of topics on various brands, products, politics and social events.
+Popular microblogging platforms like Twitter and Facebook allow users to share their opinions on a wide variety of topics.
+This popularity results in enormous amount of information covering a wide range of topics, including brands, products, politics and social events.
+As a result, sentiment analysis on such microblogging corpora has become an import area of research across the fields of NLP, machine learning, and social science \cite{o2010tweets, bollen2011twitter, bollen2009modeling}.
 
-Sentiment analysis on microblogging corpora has become an import area of research across the fields of NLP, machine learning, and social science research. 
-Sentiment analysis on microblogging has obtained special interest, because it determines the attitude of a user with respect to some topic or product and thus provides convincing information of user's interest or feedback on the topic.
+Leveraging the information contained in microblog data is non-trivial.
+Twitter, in particular, is a source of high-volume opinion data.
+On average, users post $6000$ tweets per second and $500$ million tweets per day \footnote{\url{http://www.internetlivestats.com/twitter-statistics/}}.
+As a result, the Twitter ``lexicon'' is multi-lingual and highly productive, as users post in wide variety of languages and dialects, often using slang and neologisms.
+This poses an infrastructural challenge with respect to maintaining gazetteers and other static, ``outside knowledge'' resources.
+In addition, because tweets are constrained to $140$ characters, users make creative use of acronyms, abbreviations, and emoji.
+This brevity in and of itself poses a challenge for many NLP tasks \cite{mehrotra2013improving}.
 
-For example, it may help a consumer products company get relevant feedback of public opinion on a recent product launch.
+In this paper, we propose to use bi-directional Long Short-Term Memory (LSTM) \cite{hochreiter1997long} networks which operate at {\it character-level input} to make predictions at the tweet-level.
+Such networks naturally handle the problem of very large vocabulary sizes and are able to incorporate sub-word level information into predictions.
+Our results show that character-level models outperform equivalent word-level models.
 
-In this paper, we focus on sentiment analysis on Twitter. 
-Twitter is an extremely popular microblogging platform which allows users to post messages of up to 140 characters.
-Because of the short nature of tweets, people often post twitter messages (called Tweets) frequently while attending events like product launches, movie premiers, music concerts or just to express their opinion on a trending or current topic.
-As such, they can be a valuable source of public opinion or feedback and have received considerable research interest, exemplary work among which are \cite{o2010tweets, bollen2011twitter, bollen2009modeling}.
-Realizing this importance, analyzing sentiment of Tweets has been a recurring task in the SemEval \footnote{\url{http://alt.qcri.org/semeval2016/task4/}} competitions.
+\todo{more on difference between DCNN vs LSTM -- why would we expect LSTM to be better?}
+We also compare our results with Dynamic Convolutional Neural Network (DCNN) \cite{kalchbrenner2014convolutional} which has shown state of the art performance on Twitter sentiment classification \cite{severynunitn}.
 
-%We restrict our interest on Twitter for the following reasons: 
-%People are posting on various topics on Twitter, thus it is a valuable source of data; Twitter is one of the most popular microblogging platform and the number of messages are growing everyday.
-The collected corpus can be arbitrary large.
-Users of Twitter have various background and have different language conventions.
-It is possible to collect tweets of wide diversity in both topics and language models.
+In addition, we also explore the properties of the character LSTM model and show that it learns to find meaning in composition of characters and can effectively relate sub-word information (like presence of period instead of exclamation marks) to the sentiment of the tweet.
 
-Although there exists plenty of work on text classification, some unique characteristics of tweets present special challenges for sentiment analysis:
-1) Tweets are short in length.
-There is a limitation of $140$ words for each tweet which makes analyzing them challenging \cite{mehrotra2013improving}; 2) The language used in tweets is very informal with misspellings (often intentional, like different spellings of "cool": coool, coolll, coooolll!!), new words, slangs, and URLs;
-3) The number of tweets increases at a very fast pace, and with new data comes new words, new trends in using abbreviations, which lead to a frequent problem of out-of-vocabulary words
-4) Special symbols and their combinations are often used, like emoticons and hashtags.
-
-% what we did
-Keeping in view the challenges posed by the nature of tweets, 
-in this paper, we propose to use bi-directional Long Short-Term Memory (LSTM) \cite{hochreiter1997long} 
-networks which operate at {\it character-level input} and make predictions at the Tweet-level.
-Such networks naturally handle the problem of very large vocabulary sizes and the presence of sub-word information, without having to keep many trained embedding which would be required for word-level models  with very large vocabularies.
-We find that the character-level model, interestingly, performs {\bf better} than equivalent LSTM model which operates at the word level (even when initialized with a variety of pre-trained embeddings).
-We also compare our results with Dynamic Convolutional Neural Network (DCNN) \cite{kalchbrenner2014convolutional} which has shown state of the art performance on Twitter sentiment classification \cite{severynunitn}.
-In addition, we also explore the properties of the character LSTM model and find that it learns to find meaning in composition of characters and can effectively relate sub-word information (like presence of period instead of exclamation marks) to the sentiment of the tweet.
-
-%We train our model on $1.6$ M distantly-supervised tweets collected by Go et.
-al.
-\cite{go2009twitter}, and evaluate the results on SemEval-2016 Task 4.
-Currently we focus on polarity classification, and plan to apply our model on 5-point scale classification in the future.
-
-% the structure of the paper
-This paper is laid out as follows: in Section 2, we describe prior work on sentiment classification using deep learning.
+The rest of this paper is laid out as follows: in Section 2, we describe prior work on sentiment classification using deep learning.
 In Section 3, we describe the structure of our model.
@@
 Given the length limitations on tweets, sentiment analysis of tweets is often considered similar to sentence-level sentiment analysis \cite{kouloumpis2011twitter}.
+
 However, phrase and sentence level approaches can hardly define the sentiment of some specific topics.
@@
 \cite{wang2011topic} proposed a hashtag-level sentiment classification method  to generate the overall sentiment polarity for a given hashtag.
+
 Recently, following the work of \cite{mikolov2013efficient} some researchers used neural network to implement sentiment classification.
@@
 al.
-\cite{tang2014learning} used ConvNets to learn sentiment specific word embedding (SSWE), which encodes sentiment information in the continuous
-representation of words.
+\cite{tang2014learning} used ConvNets to learn sentiment specific word embedding (SSWE), which encodes sentiment information in the continuous representation of words.
+
 Furthermore, Kalchbrenner \cite{kalchbrenner2014convolutional} proposed a Dynamic Convolutional Neural Network (DCNN) which uses dynamic k-max pooling, a global pooling operation over linear sequences.
@@
 They showed that the deep ConvNets does not require knowledge of words and thus can work for different languages.
+
 LSTM \cite{hochreiter1997long} is another state-of-the-art semantic composition models for sentiment classification \cite{li2015tree}.
@@
 Recently, there has been a surge of research interest in exploring character-input models for a variety of natural language processing tasks like: language modeling \cite{kim2015character}, POS tagging \cite{ling2015finding}, dependency parsing \cite{ballesteros2015improved} and machine translation \cite{ling2015character}.
+
 Our exploration is in a similar direction, focused on microblogging data which is naturally morphologically rich, and on using LSTM which are powerful models of sequence data and hence a natural choice for text modeling.
@@
 Note here $c$ is the convolution filter width, which is a hyperparameter.
+
 The max-pooling operations presented in \cite{kalchbrenner2014convolutional} are different from the regular max-pooling.
@@
 $k$-max pooling takes the top $k$ maximum activations as opposed to just the maximum activation and the value of $k$ is selected dynamically based on the following formula: $k_l = \max \left( k_{top}, 	\lceil \frac{L-l}{L} s\rceil \right)$, where $l$ is number of current convolution layer, $L$ is total number of convolutions and $k_{top}$ is a fixed hyperparameter.
+
 Note that while \cite{kalchbrenner2014convolutional} used multiple layers of convolutions and max-pooling, subsequent work found that using a single layer of convolution and max-pooling gives similar results \cite{kim2014convolutional} \cite{severynunitn}.
@@
 
-\begin{figure}
+\begin{figure}[t]
 	\centering
@@
 \subsection{Experiments}
+\todo{verify DCNN results, run on semeval 2016, run using our embeddings?}
+
+\todo{add latest ascii/eye results}
+
+\todo{evaluate using macro-avg F1, not accuracy}
+
+Implementations are available at: \url{https://github.com/umass-semeval/semeval16}
+
 We compare the performance of character-level models against word-level models.
@@
 \subsection{Qualitative Analysis}
+
+\todo{comparison with word-level model results}
+
 Figure ~\ref{fig:cool} shows the effect of character repetition on model confidence over the course of the sequence, where confidence is computed using the softmax function:
@@
 
-Figure ~\ref{fig:basketball} shows that our model learns the effect of negation.
+Figure ~\ref{fig:basketball} shows that our model learns the effect of negation. \todo{more details}
 

### 2015-12-20 paper/paper.tex
removed copy
minor changes

@@
 \hline
-      & \multicolumn{2}{l}{Go et. al. (1.6M)} & \multicolumn{2}{l}{SemEval2016} \\
+      & \multicolumn{2}{l}{Go et. al. (1.6M)} & \multicolumn{2}{l|}{SemEval2016} \\
       & neg               & pos               & neg            & pos            \\
@@
 \subsection{Experiments}
+Implementations are available at: \url{https://github.com/umass-semeval/semeval16}\
+
 We compare the performance of character-level models against word-level models.
@@
 Figure ~\ref{fig:basketball} shows that our model learns the effect of negation.
+The confidence in positive prediction falls sharply when after the model reads the first two characters of the negation ``not".
+Similarly, the confidence also falls after reading the suffix ``n't" which negates the meaning of the auxiliary verb.
 
@@
 
-To further analyze whether related spellings or variants of a word are mapped to nearby points in the final encoding, we fed as input individual words (from the dataset vocabulary) to the character input LSTM model and obtained the encoding of the word after the last layer (before the softmax).
-We then find the closest neighbors of every word in the encoding space.
-Table ~\ref{table:word-sim} shows words alongside their top nearest neighbors where closeness is measured by cosine similarity.
-As desired, orthographic variations of the same word ``type'' are close together in the embedding space.
+Table ~\ref{table:word-sim} shows that related spellings or variants of a word are mapped to nearby points in the final encoding.
+The word encodings are obtained by feeding as input individual words (from the dataset vocabulary) to the character input LSTM model and obtaining the encoding of the words at the last layer (before the softmax).
+We then find the closest neighbors of every word in this encoding space, where closeness is measured by cosine similarity.
+As desired, orthographic variations of the same word ``type'' are close together in the encoding space.
 In particular, it is worth noting that "woohoo" with different number of ``o'' are all close.

### 2015-12-21 paper/paper.tex
added future work

@@
 
-As Table ~\ref{table:results} shows, \cite{kalchbrenner2014convolutional} outperform our character LSTM on the Go dataset.
-However, the results presented for our character and word-level models achieve competitive performance on the Go dataset without tuning hyperparameters such as learning rate and network width.
+As Table ~\ref{table:results} shows, word/dcnn \cite{kalchbrenner2014convolutional} outperform our character LSTM on the Go dataset.
+Note that this is the published result.
+The results presented for our character and word-level models achieve competitive performance on the Go dataset without tuning hyperparameters such as size of embeddings, number of hidden units and learning rate.
+We expect the results to improve after a thorough tuning of these parameters.
 
@@
 
-Table ~\ref{table:word-sim} shows that related spellings or variants of a word are mapped to nearby points in the final encoding.
+Table x`~\ref{table:word-sim} shows that related spellings or variants of a word are mapped to nearby points in the final encoding.
 The word encodings are obtained by feeding as input individual words (from the dataset vocabulary) to the character input LSTM model and obtaining the encoding of the words at the last layer (before the softmax).
@@
 
+Future work will explore other variants on the architecture presented in this paper, for example having explicit representation of words at the higher layers composed from the character encodings of the lower layer, inspired by the results of \cite{ling2015finding}.
+Empirical evidence, as well as intuition, suggests that the character-level models need to be {\it deep} in order to get reasonable performance.
+In this work we experimented with a two layer model and would like to try deeper models where Highway networks \cite{srivastava2015training} like connections between layers could help training.
+Another direction of research to get better sentiment analysis results could be to explore the addition of additional engineered features into the models.
+
 \bibliography{paper}

### 2015-12-21 paper/paper.tex
compiled pdf

@@
 Microblogging platforms like Twitter and Facebook allow users to share their opinions on a wide variety of topics.
-This popularity results in enormous amount of information covering a wide range of topics on various brands, products, politics and social events.
+Users of these platforms generate enormous amount of information covering a wide range of topics on various brands, products, politics and social events.
 
-Sentiment analysis on microblogging corpora has become an import area of research across the fields of NLP, machine learning, and social science research. 
+Sentiment analysis on microblogging corpora has become an important area of research across the fields of NLP, machine learning, and social science research. 
 Sentiment analysis on microblogging has obtained special interest, because it determines the attitude of a user with respect to some topic or product and thus provides convincing information of user's interest or feedback on the topic.
-
 For example, it may help a consumer products company get relevant feedback of public opinion on a recent product launch.
